---
title: "Assignment 11 - Shiny"
author: "Group 8, Joseph Bannon(jb9war)"
date: "Due Date: 11:59pm, Mar 27"
output:
     html_document:
          number_sections: yes
          toc: TRUE
          toc_depth: 4
          toc_float: true
          toc_collapsed: true
          theme: journal
          code_folding: show
runtime: shiny    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
library(shiny)
library(rsconnect)
library(dplyr)
library(ggplot2)
library(foreign)
library(tidyverse)
```

# Group Homework 

- You will work with your group to complete this assignment.

- Upload your Shiny app on [ShinyApps.io](https://www.shinyapps.io).

- Submit your group's shared `.Rmd` AND your URL (https://XXXXX.shinyapps.io/XXXXX) on Collab.

- Confirm this with the following comment included in your submission text box: **"Honor Pledge: I have recreated my group submission using using the tools I have installed on my own computer"**
    
- Name the files with a group name and YOUR name for your submission.

- One of the group members will **present R codes and plots for Parts 1 and 2** in class on **Mar. 28 (Mon)**. Please e-mail the instructor if you're a presenter by 11:59pm, Mar. 27. Also, if you're a presenter, please bring your laptop so that you can share your screen on zoom for the presentation.

# Part 1



- First, get the latest data from from [https://covid19.who.int/info?openIndex=2](https://covid19.who.int/info?openIndex=2).
      
   - The file should likely be named "WHO-COVID-19-global-table-data.csv"
      
   - Don't use the data for Group Assignment 10. It may not be the most recent data. 

- Create a subset including 3 countries per `WHO region` (Africa, Americas, Eastern Mediterranean, Europe, South-East Asia, Western Pacific). You can choose any three countries within each `WHO region` to compare the mortality rate (`mutate(rate = "Deaths - cumulative total"/"Cases - cumulative total")`). The selected countries can be the same as those for Group Assignment 10.

- Create a shiny app to produce bar plots (with error bars), along with the control type of `selectInput()`.

     - You need to use `selectInput()` in the `ui` to select one level from the `WHO region` variable (with 6 levels: Africa, Americas, Eastern Mediterranean, Europe, South-East Asia, Western Pacific). 
     
     - Your output should be a barplot of mortality rates with error bars from three countries in the selected region. 

- Example: https://youmi.shinyapps.io/A11examplePart1/

## Get data


```{r}
# add your codes 
covids<- read.csv('WHO-COVID-19-global-table-data.csv')
```

```{r}
# bar graphs
covids <- cbind(Country = rownames(covids), covids)
rownames(covids) <- 1:nrow(covids)
names(covids)[2] <- 'Region'
names(covids)[9] <- 'Cumulative_deaths'
names(covids)[4] <- 'Cumulative_cases'
covid_df_region <- covids
```

## Manipulate data
```{r}
focus <- c('Brazil', 'Mexico', 'Colombia', 
           'Finland', 'Norway','Italy',
           'Indonesia', 'India', 'Thailand',
           'Republic of Korea', 'Japan', 'Viet Nam',
           'Qatar', 'Iraq', 'Morocco',
           'Zambia','Algeria','Kenya'
           )

covid_region_SE <- covid_df_region %>% filter(Country %in% focus) %>%
  mutate(rate = `Cumulative_deaths`/`Cumulative_cases`, 
         SE = sqrt(rate*(1-rate)/`Cumulative_cases`))
```

## Plot

```{r}
regions <- c("Americas", "South-East Asia", "Europe", "Western Pacific", "Eastern Mediterranean","Africa")
inputPanel(
  selectInput("region", label = "Select region:",
              choices = regions, selected = "Americas")
)

renderPlot({
 ggplot(covid_region_SE[which(covid_region_SE$Region == input$region),], aes(x=Country, y=rate)) + geom_col(fill = 'light blue') + theme() + geom_errorbar(aes(ymin=rate-1.96*SE, ymax=rate+1.96*SE), width=.2) + labs(x = 'WHO: # Reported Deaths / # Reported Cases', y = 'Mortality rate', title=paste0("Mortality rate grouped by Region: ", input$region))
})

```
# Part 2

- Use a dataset that is available in data repositories (e.g., [kaggle](https://www.kaggle.com/datasets)). But don't use the datasets you used for your previous assignments.

- Briefly describe the dataset you're using for this assignment (e.g., means to access data, context, sample, variables, etc...)
  
  
The data set is about titled 'Ferro (FOE) Stock Price from 1980 - 2022' and contains information about stock price of the Ferro Corporation. The columns are Date (the date between 1980-2020 every day), Open (the price when trading begins), High (the daily high), Low (the daily low), Close (the price when trading begins), Adj Close (closing price adjusted for cooperate actions).
   
- Select *two* continuous variables (denoted as "A" and "B") 


The two continuous variables selected are Date and High.


- Create an interactive scatterplot between "A" and "B" with a loess line using Shiny app. Please refer to lecture notes `DS3003_localreg_loess_note.pdf`, `DS3003_localreg_loess_code.rmd`, and `DS3003_localreg_loess_code.html` 

- You need to create two input variables for loess regression. 

    - One input variable is the degree with values of 0, 1, and 2.
    - The other input variable is the span value ranging from 0.2 to 0.9.
   
- You can choose appropriate control types from [Widget gallery](https://shiny.rstudio.com/gallery/widget-gallery.html).
 
```{r}
foe <- read_csv("FOE.csv")

```

     
```{r}
# add your codes 


shinyApp(
  ui = fluidPage(
   sidebarLayout(
    sidebarPanel( 
     sliderInput("span", label = "Span adjustment:",
              min = 0.2, max = 0.9, value = 0.5, step = 0.1),
      selectInput("degree", label = "Degree of Loess line:",
              choices = c(0, 1, 2), selected = 1),
    ),
    mainPanel(
     plotOutput("FOEplot"),
    )
   )
  ),
  
  server = function(input, output){
    
    output$FOEplot <- renderPlot({
    plot(High ~ Date, data = foe, cex = .4, xlab = 'Date (YYYY-MM-DD)',  ylab = 'Stock Price High (dollars)', main = 'loess regression of Date vs Stock Price High') 

x.val <- seq(min(foe$Date), max(foe$Date), length.out = 97)
out.lss2 <- loess(High ~ as.numeric(Date), data = foe, span = input$span, degree = as.numeric(input$degree))
y.pred <- predict(out.lss2, data.frame(Date = x.val))
lines(x.val, y.pred, col = 'green', lwd = 2)
    })
    
    
  }                                  
)

```

- Describe a relationship between A and B. Also, describe how a loess line changes as the degree increases and/or the span value increases.

The relationship between Date and High is non linear, possible cubic. There is a positive correlation between dates: 1980 - 2000, then a negative correlation between dates: 2000-2012, then a positive correlation again between dates: 2012 - 2020. The changes in the span for the loess line is that as span increases, the line becomes more general by staying closer to the mean of High through sections of the data. At a lower span, the line follows the High columns general increases closely. The difference between degree 0,1,2 is that high degree follows the small fluxations more closely, while low degree follows the mean of these small flucations. 

