---
title: "Assignment 11 - Shiny"
author: "Group 8, Joseph Bannon(jb9war)"
date: "Due Date: 11:59pm, Mar 27"
output:
     html_document:
          number_sections: yes
          toc: TRUE
          toc_depth: 4
          toc_float: true
          toc_collapsed: true
          theme: journal
          code_folding: show
runtime: shiny    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
library(shiny)
library(rsconnect)
library(dplyr)
library(ggplot2)
```

# Group Homework 

- You will work with your group to complete this assignment.

- Upload your Shiny app on [ShinyApps.io](https://www.shinyapps.io).

- Submit your group's shared `.Rmd` AND your URL (https://XXXXX.shinyapps.io/XXXXX) on Collab.

- Confirm this with the following comment included in your submission text box: **"Honor Pledge: I have recreated my group submission using using the tools I have installed on my own computer"**
    
- Name the files with a group name and YOUR name for your submission.

- One of the group members will **present R codes and plots for Parts 1 and 2** in class on **Mar. 28 (Mon)**. Please e-mail the instructor if you're a presenter by 11:59pm, Mar. 27. Also, if you're a presenter, please bring your laptop so that you can share your screen on zoom for the presentation.

# Part 1



- First, get the latest data from from [https://covid19.who.int/info?openIndex=2](https://covid19.who.int/info?openIndex=2).
      
   - The file should likely be named "WHO-COVID-19-global-table-data.csv"
      
   - Don't use the data for Group Assignment 10. It may not be the most recent data. 

- Create a subset including 3 countries per `WHO region` (Africa, Americas, Eastern Mediterranean, Europe, South-East Asia, Western Pacific). You can choose any three countries within each `WHO region` to compare the mortality rate (`mutate(rate = "Deaths - cumulative total"/"Cases - cumulative total")`). The selected countries can be the same as those for Group Assignment 10.

- Create a shiny app to produce bar plots (with error bars), along with the control type of `selectInput()`.

     - You need to use `selectInput()` in the `ui` to select one level from the `WHO region` variable (with 6 levels: Africa, Americas, Eastern Mediterranean, Europe, South-East Asia, Western Pacific). 
     
     - Your output should be a barplot of mortality rates with error bars from three countries in the selected region. 

- Example: https://youmi.shinyapps.io/A11examplePart1/

## Get data


```{r}
# add your codes 
covids<- read.csv('WHO-COVID-19-global-table-data.csv')
```

```{r}
# bar graphs
covids <- cbind(Country = rownames(covids), covids)
rownames(covids) <- 1:nrow(covids)
names(covids)[2] <- 'Region'
names(covids)[9] <- 'Cumulative_deaths'
names(covids)[4] <- 'Cumulative_cases'
covid_df_region <- covids
```

## Manipulate data
```{r}
focus <- c('Brazil', 'Mexico', 'Colombia', 
           'Finland', 'Norway','Italy',
           'Indonesia', 'India', 'Thailand',
           'Republic of Korea', 'Japan', 'Viet Nam',
           'Qatar', 'Iraq', 'Morocco',
           'Zambia','Algeria','Kenya'
           )

covid_region_SE <- covid_df_region %>% filter(Country %in% focus) %>%
  mutate(rate = `Cumulative_deaths`/`Cumulative_cases`, 
         SE = sqrt(rate*(1-rate)/`Cumulative_cases`))
```

## Plot

```{r}
regions <- c("Americas", "South-East Asia", "Europe", "Western Pacific", "Eastern Mediterranean","Africa")
inputPanel(
  selectInput("region", label = "Select region:",
              choices = regions, selected = "Americas")
)

renderPlot({
 ggplot(covid_region_SE[which(covid_region_SE$Region == input$region),], aes(x=Country, y=rate)) + geom_col(fill = 'light blue') + theme(axis.text.x=element_text(angle=90, hjust=1), text = element_text(size = 20) ) + geom_errorbar(aes(ymin=rate-1.96*SE, ymax=rate+1.96*SE), width=.2) + labs(x = 'WHO: # Reported Deaths / # Reported Cases', y = 'Mortality rate', title=paste0("Mortality rate of countries grouped by Region: ", input$region))
})

```
# Part 2

- Use a dataset that is available in data repositories (e.g., [kaggle](https://www.kaggle.com/datasets)). But don't use the datasets you used for your previous assignments.

- Briefly describe the dataset you're using for this assignment (e.g., means to access data, context, sample, variables, etc...)
  
   + *describe your data.*
   
- Select *two* continuous variables (denoted as "A" and "B") 

- Create an interactive scatterplot between "A" and "B" with a loess line using Shiny app. Please refer to lecture notes `DS3003_localreg_loess_note.pdf`, `DS3003_localreg_loess_code.rmd`, and `DS3003_localreg_loess_code.html` 

- You need to create two input variables for loess regression. 

    - One input variable is the degree with values of 0, 1, and 2.
    - The other input variable is the span value ranging from 0.2 to 0.9.
   
- You can choose appropriate control types from [Widget gallery](https://shiny.rstudio.com/gallery/widget-gallery.html).
     
```{r}
# add your codes 
```

- Describe a relationship between A and B. Also, describe how a loess line changes as the degree increases and/or the span value increases.

