---
title: "Assignment 13 - Shiny and Maps (4 pts)"
author: "Group 8, Joseph Bannon (jb9war)"
date: "Due Date: 11:59pm, Apr 4"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
runtime: shiny    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
library(tidyverse)
library(shiny)
library(shinythemes)
```

# Group Homework 

- You will work with your group to complete this assignment.

- Upload your Shiny app on [ShinyApps.io](https://www.shinyapps.io).

- Submit your group's shared `.Rmd` AND your URL (https://XXXXX.shinyapps.io/XXXXX) on Collab.

- Confirm this with the following comment included in your submission text box: **"Honor Pledge: I have recreated my group submission using using the tools I have installed on my own computer"**
    
- Name the files with a group name and YOUR name for your submission.

- One of the group members will **present R codes and plots for Parts 1 and 2** in class on **Apr. 4 (Mon)**. Please e-mail the instructor if you're a presenter by 11:59pm, Apr. 3. Also, if you're a presenter, please bring your laptop so that you can share your screen on zoom for the presentation.

# Part 1

- Get the latest data from from [https://covid19.who.int/info?openIndex=2](https://covid19.who.int/info?openIndex=2).
      
   - The file should likely be named "WHO-COVID-19-global-table-data.csv"
      
   - Don't use the data for the previous group assignments. It is not the most recent data. 

- Choose three WHO regions of interest. Note that we have 6 main regions: Africa, Americas, Eastern Mediterranean, Europe, South-East Asia, and Western Pacific.

- Create a subset including 3 countries per `WHO region`. You can choose any three countries within each `WHO region` to compare the mortality rate (`mutate(rate = "Deaths - cumulative total"/"Cases - cumulative total")`). You will have 9 countries (3 countries * 3 WHO regions).

- Using `navbarPage()`, create a shiny dashboard that contains 3 tabs where each tab has a barplot of mortality rates with error bars from three countries in the selected region. For example,

```{r}
# add your codes 
covids <- read.csv('WHO-COVID-19-global-table-data.csv')
```

```{r}
# bar graphs
covids <- cbind(Country = rownames(covids), covids)
rownames(covids) <- 1:nrow(covids)
names(covids)[2] <- 'Region'
names(covids)[9] <- 'Cumulative_deaths'
names(covids)[4] <- 'Cumulative_cases'
covid_df_region <- covids
```

```{r}
focus <- c('Brazil', 'Mexico', 'Colombia', 
           'Finland', 'Norway','Italy',
           'Indonesia', 'India', 'Thailand'
           )


covid_region_SE <- covid_df_region %>% filter(Country %in% focus) %>%
  mutate(rate = `Cumulative_deaths`/`Cumulative_cases`, 
         SE = sqrt(rate*(1-rate)/`Cumulative_cases`))


covid_ame <- covid_region_SE[which(covid_region_SE$Region == "Americas"),]
covid_euro <- covid_region_SE[which(covid_region_SE$Region == "Europe"),]
covid_sea <- covid_region_SE[which(covid_region_SE$Region == "South-East Asia"),]


```

```{r}
ui <-  navbarPage(title = "Mortality Rate", theme = shinytheme("flatly"),
            tabPanel("Americas", 
                sidebarLayout(
                    sidebarPanel(
                      checkboxInput("err_1", label = "Error bars", value = FALSE)
                    ),
                    mainPanel(
                      plotOutput("ameplot")
                    )
                )
            ),
            tabPanel("Europe",
                sidebarLayout(
                    sidebarPanel(
                        radioButtons("err_2", label = h3("Radio buttons"),
                        choices = list("No Error Bars" = FALSE, "Error Bars" = TRUE), 
                        selected = FALSE),
                    ),
                    mainPanel(
                      plotOutput("europlot")
                    )
                )         
            ), 
            tabPanel("South-East Asia",
                sidebarLayout(
                    sidebarPanel(
                      selectInput("err_3", label = h3("Select box"), 
                      choices = list("No Error Bars" = FALSE, "Error Bars" = TRUE), 
                      selected = FALSE)
                    ),
                    mainPanel(
                      plotOutput("seaplot")
                    )
                )
            )
) 
```      


- Create separate `tabPanel()` for each level of`WHO region` variable.

- Under each tab, create a bar plot (with error bars) from three countries in the selected region with a controller for including/excluding errors bars. The default plot has no error bars.

- Use different control types under each tab. Check available control types from [Widget gallery](https://shiny.rstudio.com/gallery/widget-gallery.html).
  
- Change a theme from the `shinythemes` package. See options from https://rstudio.github.io/shinythemes/.   

- Tips for  `renderPlot()`:

- Example: https://youmi.shinyapps.io/A13part1/

```{r}
# add your codes 
server <- function(input, output){
    
    output$ameplot = renderPlot({
     
     p1 <-ggplot(covid_ame, aes(x=Country, y=rate)) + geom_col(fill = 'light blue') + theme() + labs(x = 'WHO: # Reported Deaths / # Reported Cases', y = 'Mortality rate', title="Mortality rate grouped by Region: Americas", )
     print(p1)
      
     if (input$err_1) { # modify this line
      p1_e <- p1 + geom_errorbar(aes(ymin=rate-1.96*SE, ymax=rate+1.96*SE), width=.2)  # add error bars
      print(p1_e)
     }})
    
    output$europlot = renderPlot({
     
     p2 <-ggplot(covid_euro, aes(x=Country, y=rate)) + geom_col(fill = 'light blue') + theme() + labs(x = 'WHO: # Reported Deaths / # Reported Cases', y = 'Mortality rate', title="Mortality rate grouped by Region: Europe", )
     print(p2)
      
     if (input$err_2) { # modify this line
      p2_e <- p2 + geom_errorbar(aes(ymin=rate-1.96*SE, ymax=rate+1.96*SE), width=.2)  # add error bars
      print(p2_e)
     }})
    
    output$seaplot = renderPlot({
  
     p3 <-ggplot(covid_sea, aes(x=Country, y=rate)) + geom_col(fill = 'light blue') + theme() + labs(x = 'WHO: # Reported Deaths / # Reported Cases', y = 'Mortality rate', title="Mortality rate grouped by Region: Europe", )
     print(p3)
      
     if (input$err_3) { # modify this line
      p3_e <- p3 + geom_errorbar(aes(ymin=rate-1.96*SE, ymax=rate+1.96*SE), width=.2)  # add error bars
      print(p3_e)
     }})
  }
```

```{r}
shinyApp(ui = ui, server = server, options = list(height = 600))
```


# Part 2

- Create a choropleth US map like the Trump Impeachment example. You need to add any state-related variable (e.g., the number of universities per state, the crime rate per state) onto the US map.

- Use a dataset that is available in data repositories (e.g., [kaggle](https://www.kaggle.com/datasets)). But don't use the built-in datasets or datasets from packages. You need to find a state-related variable online. 

- Briefly describe the dataset you're using for this assignment (e.g., means to access data, context, sample, variables, etc...)
  
   + *describe your data.*
   
- Create a choropleth map that displays data on a map by shading states with different colors.

- You can use any map-related functions we've learned in class or any other available functions for creating maps. Your map can be a static map or an interactive map.

```{r}
# add your codes 
```

- Describe your choropleth map.

